Smart Spreading Factor Assignment for LoRaWANs

### 2019-02-27 paper.tex
prediction results explanation

It is possible to generate mass amount of LoRaWAN transmission logs for different topologies using our simulator. Dataset size is directly proportional to simulation duration. Thus, increasing the simulation duration, improves the prediction accuracy up to some extent. In real world deployments, NS can keep track of transmissions and it can create a classifier daily basis, then gateways can request from nodes to use suggested SFs.

There are three features in the dataset generated by the simulation tool. Dataset features are: X and Y coordinates of the transmission source, SF of the transmission. X and Y coordinate feature values are continuous numbers, SF feature values are integer numbers between 7 to 12. Dataset class label is the result of the transmission. Dataset class label values are successful, interfered and under sensitivity. DTC and SVM prediction schemes are integrated into the simulation tool to study smart SF assignment schemes. A classifier is trained from generated dataset and this classifier is used for selecting optimum SF for the nodes in the network.

The tool first runs a simulation with random SF scheme. After random SF scheme simulation completed, transmission logs are combined into three feature columns and one class label column to create the dataset. Then the dataset is fed to Python scikit-learn DTC or SVM classifiers for training phase. After the classifier model is built, second simulation is run with the trained classifier. The tool selects optimum SF for transmissions considering prediction of the transmissions. For every transmission, the classifier calculates the transmission result for the lowest possible SF. If the transmission result is classified as interfered, then the tool increases the SF and predicts a new transmission result. If the new transmission result is classified as successful, then simulator continue to execute with selected SF. If no SF transmission result is predicted as successful, then the tool selects the lowest possible SF and continue to execute.

@@

\subsection{Single Gateway}
In Figure \ref{fig:sf_pdr}, PDR plot of various SF assignment is shown. Randomly generated network topology radius is set to 3000 meters and number of gateways is set to 1. Increasing SFs increases air time. It increases the number of collisions thus decreases the PDR. High SF schemes gives poor PDR results as the number of nodes increases. Since network topology radius is quite small, all SFs can reach to gateway.

@@

In Figure \ref{fig:r_pdr}, PDR plot of various network radius is shown. Number of gateway is set to 1 and lowest SF assignment scheme is used. Increasing network radius, increases the number of under sensitivity transmissions thus decreases the PDR of network.

@@

\subsection{Multiple Gateway}
In Figure \ref{fig:gw_pdr}, PDR plot for various number of gateways is shown. Randomly generated topology radius is set to 3000 meters and the lowest SF assignment scheme is used. Increasing number of gateways, decreases the SFs of nodes and decreases air time. This decreases number of collisions, hence, increases the PDR of network when network radius is constant.

@@

\subsection{Smart Spreading Factor Schemes}
PDR plot of the lowest SF scheme and the smart prediction schemes are shown in Figure \ref{fig:prediction_pdr}. Randomly generated network radius is set to 5000 meters and number of gateways is set to 3. Prediction model needs nodes' locations and three gateways are enough to locate position of nodes by triangulation. In Table \ref{table:prediction_pdr}, PDR values for various network radii are presented. Both smart SVM and smart DTC schemes give better PDR than lowest SF schemes when number of nodes increases. Increasing number of nodes, increases number of interferences. Smart schemes improve network performance when LoRa interference is high. Moreover, smart schemes give better results when nodes are deployed closer to the gateway, since nodes have margin to increase their SFs when they are deployed closer to the gateway. If a node is far away from gateway, then smart schemes cannot increase the SF to avoid interference since the assigned SF is already high.

In Table \ref{table:prediction_accuracy}, prediction accuracy of smart SVM and smart DTC schemes are presented for various network radii and number of nodes. Prediction accuracy is not directly proportional to network PDR. Correct prediction of an interfered transmission may not increase the PDR but increases the accuracy. Smart DTC gives better PDR results than smart SVM, even overall SVM prediction accuracy is high. LoRaWAN simulation usually produces imbalanced dataset. Under sensitivity transmission labeled data points are less than successful transmission labeled data points. Interfered transmission labeled data points are even less then Under sensitivity transmission labeled data points. DTC gives better accuracy for interfered transmission label. Correct classification of interfered transmissions yields better network PDR results for smart SF scheme.

@@

\section{Conclusion} \label{Conclusion}
In this paper, after a brief introduction about LPWAN technologies, LoRa modulation basics and spreading factor assignment issue is discussed. We present a discrete event simulator which is developed from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes. Moreover, we show how same SF collisions can be avoided, hence, we propose a machine learning based solutions called smart DTC scheme and smart SVM scheme. We present simulation results for the lowest SF assignment scheme and the proposed schemes. Simulation results show that, the proposed smart schemes can increase network performance for LoRaWAN networks, especially, when the nodes are deployed close to gateways.

>>>>

It is possible to generate mass amount of LoRaWAN transmission logs for different topologies using our simulator. Training dataset size is directly proportional to simulation duration. Thus, increasing the simulation duration, improves the prediction accuracy up to some extent. In real world deployments, NS can keep track of transmissions and it can create a classifier daily basis, then gateways can request from nodes to use suggested SFs.

There are three features in the dataset generated by the simulation tool. Features of the dataset are: X coordinate of the transmission source, Y coordinate of the transmission source and SF of the transmission. X and Y coordinate feature values are continuous numbers. SF feature values are integer numbers between 7 to 12. Class label of the dataset is the result of the transmission. Class label values are successful, interfered and under sensitivity. DTC and SVM prediction schemes are integrated into the simulation tool to study smart SF assignment schemes. A classifier is trained from generated dataset and this classifier is used for selecting optimum SF for the nodes in the network.

The tool first runs a simulation with random SF scheme. After random SF scheme simulation completed, transmission logs are combined into three feature columns and one class label column to create the training dataset. Then the dataset is fed to Python scikit-learn DTC or SVM classifiers for training phase. After the classifier model is built, second simulation is run with the trained classifier. The tool selects optimum SF for transmissions considering prediction of the transmissions. For every transmission, the classifier predicts the transmission result for the lowest possible SF. If the transmission result is predicted as interfered, then the tool increases the SF and predicts a new transmission result. If the new transmission result is classified as successful, then simulator continue to execute with selected SF. If no SF transmission result is predicted as successful, then the tool selects the lowest possible SF and continue to execute.

@@

\subsection{Single Gateway}
In Figure \ref{fig:sf_pdr}, PDR plots of various SF assignment schemes are shown. Randomly generated network topology radius is set to 3000 meters and number of gateways is set to 1. Increasing SFs increases air time. This increases the number of collisions thus decreases the PDR. High SF schemes gives poor PDR results as the number of nodes increases. Since network topology radius is quite small, all SFs can reach to gateway.

@@

In Figure \ref{fig:r_pdr}, PDR plots of various network radii are shown. Number of gateway is set to 1 and lowest SF assignment scheme is used. Increasing network radius, increases the number of under sensitivity transmissions thus decreases the PDR of network.

@@

\subsection{Multiple Gateway}
In Figure \ref{fig:gw_pdr}, PDR plots for various number of gateways are shown. Randomly generated topology radius is set to 3000 meters and the lowest SF assignment scheme is used. Increasing number of gateways, decreases the SFs of nodes and decreases air time. This decreases number of collisions, hence, increases the PDR of network when network radius is constant.

@@

\subsection{Smart SF Schemes}
PDR plots of the lowest SF, random SF and the smart prediction schemes are shown in Figure \ref{fig:prediction_pdr}. Randomly generated network radius is set to 5000 meters and number of gateways is set to 3. Prediction model needs nodes' locations and three gateways are enough to locate position of nodes by triangulation. In Table \ref{table:prediction_pdr}, PDR values for various network radii are presented. Both smart SVM and smart DTC schemes give better PDR than lowest SF schemes when number of nodes increases. Increasing number of nodes, increases number of interferences. Smart schemes improve network performance when LoRa interference is high. Moreover, smart schemes give better results when nodes are deployed closer to the gateway, since nodes have margin to increase their SFs when they are deployed closer to the gateway. If a node is far away from gateway, then smart schemes cannot increase the SF to avoid interference since the assigned SF is already high.

In Table \ref{table:prediction_accuracy}, prediction accuracy of smart SVM and smart DTC schemes are presented for various network radii and number of nodes. Prediction accuracy is not directly proportional to network PDR. Correct prediction of an interfered transmission may not increase the PDR but increases the prediction accuracy. Smart DTC gives better PDR results than smart SVM, even overall SVM prediction accuracy is higher. LoRaWAN simulation usually produces imbalanced training dataset. Number of under sensitivity transmission labeled data points are less than number of successful transmission labeled data points. Number of interfered transmission labeled data points are even less then number of under sensitivity transmission labeled data points. In our case, DTC gives better prediction accuracy for interfered transmission labeled data points. Correct classification of interfered transmissions yields better network PDR results, thus smart DTC scheme gives better network PDR results.

@@

\section{Conclusion} \label{Conclusion}
In this paper, after a brief introduction about LPWAN technologies, LoRa modulation basics and spreading factor assignment issue is discussed. We present a open source discrete event simulator which is developed from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes. Moreover, we show how same SF collisions can be avoided, hence, we propose a machine learning based solutions called smart DTC scheme and smart SVM scheme. We present simulation results for the lowest SF assignment scheme and the proposed schemes. Simulation results show that, the proposed smart schemes can increase network performance for LoRaWAN networks, especially, when the nodes are deployed close to gateways.
